Financialisation of Commodity Markets \\\large{Co-movement Behind-the-Scenes}

### 2025-07-04 paper.tex
Deploying to gh-pages from @ bautheac/financialisation-co-movement@342e653499d97540ee93e29b1a6bb8dbd4a5f3c9 🚀

Table~\ref{tbl-stats-combined} displays descriptive statistics for the
commodity series considered in the study; the results show a clear

@@

(XLME)---show a lower figure with five og them significant---Crude

@@

\textbf{CHP} is low and vice et versa. Table~\ref{tbl-stats-combined}
further shows the mean returns and volatilities for high
(\textbf{contango}) and low (\textbf{backwardation}) \textbf{CHP}
regimes for each time period while
Table~\ref{tbl-regime-diference-tests} shows the results of difference
tests for returns mean and variance between the two regimes.\\

@@

Financialisation was a global phenomenon, through the channel of index
investment it simultaneously influenced the entire spectrum of liquid
commodities, eventually spurring the transformation of commodity markets
from a collection of idiosyncratic markets into a more cohesive
collection akin to an asset class. In this context, interconnectedness
emerges as a crucial aspect that warrants thorough examination.\\
We strive to address this issue by first examining the average returns
pairwise correlations between commodities across the entire
cross-section of the commodity assets considered in the study as well as
by country, sector and sub-sector.
Table~\ref{tbl-correlations-inner-periods} shows the results for the
four periods of interest with the results reported for the entire period
as well as for periods of \textbf{backwardation} and \textbf{contango}.
For the entire sample the average pairwise correlation jumps from 0.07
in the first period to 0.16 over financialisation. This finding
indicates higher interdependence within the full cross-section of
commodities starting with the financialisation period, consistent with
the empirical findings in \citet{fryMcKibbin_evolution_2023} and
\citet{mayer_financialization_2017} as well as with the theoretical
model of \citet{basak_model_2016}. The sector analysis sheds light on
the driving force behind these changes. The US metals see the sharpest
increase in pairwise correlations going from 0.22 over the past period
to 0.56 over financialisation while the three other sectors see a much
smaller percentage increase during financialisation relative to the past
period. These results thus suggest that the US metals sector was the
driving force behind increased average pairwise correlations during
financialisation, a new observation to the best of our knowledge. The
pattern of correlations is very similar across phases of
\textbf{backwardation} and \textbf{contango} for all the sectors and
sub-sectors over the past and financialisation periods. In this context,
the ontological consequences of financialisation are better observed
from an investment perspective as illustrated in
Table~\ref{tbl-correlations-cross-periods} where we observe average
pairwise correlation coefficients between equally weighted portfolios of
US traded commodity assets grouped by sub-sector. Although modest in
magnitude, these exhibit a telling pattern, mirroring that shown by the
individual commodity returns in Table~\ref{tbl-stats-combined}, with a
substantially higher average coefficient over \textbf{backwardation}
phases in the first period (0.12 vs.~0.03) and the pattern reversed over
financilisation with an average coefficient substantially higher in
\textbf{contango} (0.12 vs.~0.02).\\
We break down Table~\ref{tbl-correlations-inner-periods}'s average
correlations by year in Table~\ref{tbl-correlations-inner-years}. For
the full sample of assets, the average pairwise correlations remain in
the 0.06--0.08 range from 1998 through 2003, before suddenly increasing
to 0.12 from 2003 to 2004. This suggests 2004 as a distinct break point
for the onset of financialisation, as noted in several studies
\citep{baker_financialization_2021, tang_index_2012}. We further observe
that the US metals appear to have played a major role in this as the
increase seems driven by the sector which average pairwise correlation
more than doubles (0.49 vs 0.22) over the period.

@@

\textbf{backwardation}. The onset of financialisation leads to a
substantial improvement in the performance of the CHP factor with its
average \(R^{2}\), albeit still small at 5.4\%, almost tripling from the
past to the financialisation period in contrast to the other factors
that show weaker dynamics with a slight decrease and a much smaller
increase for the open interest and term structure factors respectively.
Here again the increase in pricing performance for the CHP factor is
driven by the metals sector, particularly the precious sub-sector with
the average \(R^{2}\) for the group jumping to almost 26\% during
financialisation from 1.2\% over the past period. The corresponding
figures of 22\% and 1.1\% for the entire US metals sector stand in sharp
contrast with the agricultural sector, which exhibits no notable change,
and the energy sector, where pricing performance drops substantially
from 9.2\% down to 1.6\%. The improvement in the CHP factor's pricing
performance remarkably comes predominantly from the long leg (low CHP
assets: Keynesian ``normal backwardation'') of the factor\footnote{Results
  available from the authors upon request.} with the corresponding
mimicking portfolio's performance itself driven by the metals, the
precious sub-sector in particular. Interestingly, although the term
structure factor itself only sees a very modest increase, its short leg
(low roll-yield assets: Working's backwardation) shows a sizeable rise
in pricing performance, yet again driven the metals sector. The pattern
of results for both factors extends to the GB traded assets and suggest
a convergence of Keneysian and Working's term structure perspective over
the financialisation period.\\

@@

The crisis period (late 2008 to mid-2013) shows a contrasting pattern of
results. Returns fall sharply relative to the financialisation period
(Table~\ref{tbl-stats-combined}) with eighteen US commodities along with
all six GB traded metals showing lower mean returns. Returns for equally
weighted portfolios of both sets of assets fall down to 9.4\% and 5.3\%
from 15.9\% and 21.8\% respectively. Returns are mostly higher during
phases of \textbf{backwardation} relative to \textbf{contango} with
fifteen US assets and five of the GB traded metals showing higher mean
returns in \textbf{backwardation} and the pattern extending to equally
weighted portfolios of the two sets of assets. Volatility is moderately
higher for both sets with fifteen US commodities and three GB metals
showing higher figures and interestingly eminently higher over phases of
\textbf{contango}; out of the US complex only Sugar-\#11 (IFUS) and none
of the GB traded metals show higher volatility figures over
\textbf{backwardation} phases. The symmetry between the dynamics of CHP
and \textbf{CHP}'s explanatory power on individual commodity returns
breaks (Table~\ref{tbl-time-reg-US-CHP}) with both now seemingly
converging; the US cross-sectional average \(R^{2}\) at the individual
level bouncing back up (22.5\% vs 20.8\%) while that at the aggregate
level rises further (14.7\% vs 7.1\%).

@@

The returns pull-back continues into the post-crisis period
(Table~\ref{tbl-stats-combined}) with twenty-two US and four GB traded
assets showing lower mean returns relative to the crisis while the same
number of US commodities and all GB metals show lower figures relative
to the financialisation period; mean returns for equally weighted
portfolios of US and GB assets drop down to -0.2\% and 2.7\%
respectively. Interestingly volatility also recedes substantially with
twenty US assets and all GB metals showing lower volatility figures
relative to both the crisis and financialisation periods suggesting that
investment outflows from the commodity futures complex have had a lower
impact on volatility than the corresponding inflows witnessed over the
financialisation period. The figures for the equally weighted portfolios
of US and GB assets drop to 9.9\% and 16.2\% respectively, very close to
pre-financialisation levels.\\
The pattern of reversion toward the pre-financialisation Keynesian
paradigm initiated during the crisis continues stronger into the
post-crisis period with mean returns higher during phases of
\textbf{backwardation} relative to \textbf{contango}, more so than over
the crisis period with seventeen US commodities showing higher mean
returns over \textbf{backwardation} phases. The difference
(Table~\ref{tbl-regime-diference-tests}) is significant at the 5\% level
for Lumber (XCME) and at the 10\% level for Palladium (XNYM) while no
difference is significant for the assets showing higher figures over
phases of \textbf{contango}. The pattern is more pronounced for the GB
traded subset with all six assets showing higher mean returns in
\textbf{backwardation} with the difference significant at the 1\% level
for Nickel-primary (XLME), 5\% level for Aluminium (XLME) and 10\% level
for Zinc (XLME). It also extends to equally weighted portfolios of US
and GB traded assets that both show higher mean returns during
\textbf{backwardation} with the difference significant at the 5\% level
for the GB portfolio and the figure close to the 10\% level significance
threshold for the portfolio of US assets. Volatilities are overall
higher in \textbf{contango} although less so than over the previous
period with seventeen US commodities showing higher figures over
\textbf{contango} against twenty-three in the crisis period. The pattern
extends to three of the six GB traded metals with the difference
significant at the 1\% level for Copper (XLME) and Tin (XLME) and at the
5\% level for Zinc (XLME) while differences are not significant for
Aluminium (XLME), Lead (XLME) and Nickel-primary (XLME) that exhibit
higher volatility over \textbf{backwardation} phases. Equally weighted
portfolios of US and GB traded assets on the other hand both show higher
volatility in \textbf{contango} with the difference significant at the
1\% level for both. Over the financialisation period in contrast
volatilities were predominantly higher in \textbf{backwardation} for US
traded assets with fourteen commodities showing higher volatility
figures in these phases; the post-crisis pattern of results for
volatilities thus likewise shows signs of reversion to
pre-financialisation standards. After showing signs of convergence in
the crisis period the dynamics of the explanatory power of CHP and
\textbf{CHP} part ways (Table~\ref{tbl-time-reg-US-CHP}), seemingly
reverting to the Keynesian paradigm observed pre-financialisation,
further suggesting a return to legacy commodity specific dynamics. The
recovery initiated in the crisis period for that of CHP is confirmed
with a stronger increase, 27.2\% vs.~22.5\%, strongest for the metals
sector, 37.3\% vs.~23.7\%, while the energy sector shows the opposite
pattern: 11\% vs.~16\%. In contrast, after increasing stronger in the
crisis, that for \textbf{CHP} falls overall post-crisis, 7.5\% vs
14.7\%, with the decrease strongest for the energy sector in a relative
sense: 2.5\% vs.~14.8\%.

@@

\caption{\label{tbl-stats-combined}This table shows mean returns and
corresponding volatility for the twenty four individual US commodities
and the six GB metals considered in the study as well as for two equally
weighted portfolios formed from the US commodities and the GB metals
respectively across the four periods of interest (past: 1997-2003;
financialisation: 2004-2008; crisis: 2008-2013; post-crisis: 2013-2018).
for each period the figures are shown independently for the entire
period as well as for phases of aggregate backwardation (aggregate CHP
\(\leq\) period median) and aggregate contango (aggregate CHP
\textgreater{} period median). Mean values significant at the 1\%, 5\%
and 10\% level are marked with ***, ** and * respectively. Aggregate CHP
construction and corresponding regime definitions are discussed in
Section~\ref{sec-methods} while the results are discussed in
Section~\ref{sec-results}.}

@@

\textcolor{black}{\textbf{asset}} & \textcolor{black}{\textbf{regime}} & \textcolor{black}{\textbf{estimate}} & \textcolor{black}{\textbf{past}} & \textcolor{black}{\textbf{financialisation}} & \textcolor{black}{\textbf{crisis}} & \textcolor{black}{\textbf{post-crisis}}\\

@@

\textbf{Corn-\#2 yellow (XCBT)} & \textbf{whole period} & \textbf{mean} & \textcolor[HTML]{4285f4}{3.48\%} & \textcolor[HTML]{4285f4}{21.37\%} & \textcolor[HTML]{4285f4}{6.89\%} & \textcolor[HTML]{4285f4}{-4.51\%}\\
\textbf{} & \textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{23\%} & \textcolor[HTML]{4285f4}{29.54\%} & \textcolor[HTML]{4285f4}{36.15\%} & \textcolor[HTML]{4285f4}{21.8\%}\\
\textbf{} & \textbf{backwardation} & \textbf{mean} & \textcolor[HTML]{4285f4}{-4.86\%} & \textcolor[HTML]{4285f4}{19.26\%} & \textcolor[HTML]{4285f4}{20.48\%} & \textcolor[HTML]{4285f4}{-8.74\%}\\
\textbf{} & \textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{23.07\%} & \textcolor[HTML]{4285f4}{29.14\%} & \textcolor[HTML]{4285f4}{31.59\%} & \textcolor[HTML]{4285f4}{21\%}\\
\textbf{} & \textbf{contango} & \textbf{mean} & \textcolor[HTML]{4285f4}{11.3\%} & \textcolor[HTML]{4285f4}{23.54\%} & \textcolor[HTML]{4285f4}{-4.43\%} & \textcolor[HTML]{4285f4}{0.78\%}\\

@@

\textbf{} & \textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{23\%} & \textcolor[HTML]{4285f4}{30.01\%} & \textcolor[HTML]{4285f4}{40.14\%} & \textcolor[HTML]{4285f4}{22.68\%}\\
\textbf{Oats (XCBT)} & \textbf{whole period} & \textbf{mean} & \textcolor[HTML]{4285f4}{5.49\%} & \textcolor[HTML]{4285f4}{21.62\%} & \textcolor[HTML]{4285f4}{8.89\%} & \textcolor[HTML]{4285f4}{-0.87\%}\\
\textbf{} & \textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{32.22\%} & \textcolor[HTML]{4285f4}{31.04\%} & \textcolor[HTML]{4285f4}{34.67\%} & \textcolor[HTML]{4285f4}{29.5\%}\\
\textbf{} & \textbf{backwardation} & \textbf{mean} & \textcolor[HTML]{4285f4}{-5.39\%} & \textcolor[HTML]{4285f4}{7.82\%} & \textcolor[HTML]{4285f4}{7.85\%} & \textcolor[HTML]{4285f4}{2.56\%}\\
\textbf{} & \textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{29.02\%} & \textcolor[HTML]{4285f4}{33.82\%} & \textcolor[HTML]{4285f4}{31.43\%} & \textcolor[HTML]{4285f4}{31.55\%}\\

@@

\textbf{} & \textbf{contango} & \textbf{mean} & \textcolor[HTML]{4285f4}{16.95\%} & \textcolor[HTML]{4285f4}{*35.38\%} & \textcolor[HTML]{4285f4}{11.68\%} & \textcolor[HTML]{4285f4}{-3.17\%}\\
\textbf{} & \textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{35.22\%} & \textcolor[HTML]{4285f4}{28.1\%} & \textcolor[HTML]{4285f4}{37.7\%} & \textcolor[HTML]{4285f4}{27.5\%}\\
\textbf{Soybean meal (XCBT)} & \textbf{whole period} & \textbf{mean} & \textcolor[HTML]{4285f4}{7.29\%} & \textcolor[HTML]{4285f4}{12.06\%} & \textcolor[HTML]{4285f4}{8.12\%} & \textcolor[HTML]{4285f4}{-0.33\%}\\
\textbf{} & \textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{26.34\%} & \textcolor[HTML]{4285f4}{32.55\%} & \textcolor[HTML]{4285f4}{31.78\%} & \textcolor[HTML]{4285f4}{25.11\%}\\
\textbf{} & \textbf{backwardation} & \textbf{mean} & \textcolor[HTML]{4285f4}{13.15\%} & \textcolor[HTML]{4285f4}{8.22\%} & \textcolor[HTML]{4285f4}{16.04\%} & \textcolor[HTML]{4285f4}{-3.52\%}\\

@@

\textbf{} & \textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{27.77\%} & \textcolor[HTML]{4285f4}{33.06\%} & \textcolor[HTML]{4285f4}{25.7\%} & \textcolor[HTML]{4285f4}{25.1\%}\\
\textbf{} & \textbf{contango} & \textbf{mean} & \textcolor[HTML]{4285f4}{1.52\%} & \textcolor[HTML]{4285f4}{15.91\%} & \textcolor[HTML]{4285f4}{2.32\%} & \textcolor[HTML]{4285f4}{2.22\%}\\
\textbf{} & \textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{24.94\%} & \textcolor[HTML]{4285f4}{32.13\%} & \textcolor[HTML]{4285f4}{36.85\%} & \textcolor[HTML]{4285f4}{25.32\%}\\
\textbf{Soybean oil (XCBT)} & \textbf{whole period} & \textbf{mean} & \textcolor[HTML]{4285f4}{6.21\%} & \textcolor[HTML]{4285f4}{15.57\%} & \textcolor[HTML]{4285f4}{4.04\%} & \textcolor[HTML]{4285f4}{-7.39\%}\\
\textbf{} & \textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{21.43\%} & \textcolor[HTML]{4285f4}{28.35\%} & \textcolor[HTML]{4285f4}{25.75\%} & \textcolor[HTML]{4285f4}{18.95\%}\\

>>>>

Table~\ref{tbl-stats-regime-tests} displays descriptive statistics for
the commodity series considered in the study; the results show a clear

@@

(XLME)---show a lower figure with five of them significant---Crude

@@

\textbf{CHP} is low and vice et versa.
Table~\ref{tbl-stats-regime-tests} further reports the dominating regime
for each asset-period-moment combination along with the corresponding
results of difference tests for returns mean and variance between the
two regimes\footnote{The estimate values per regime as well as the
  p-values for the corresponding difference tests are available in the
  attached
  \href{http://bautheac.duckdns.org:3838/bautheac/co-movement/}{Internet
  Appendix}.}.\\

@@

Financialisation was a global phenomenon, through the channel of index
investment it simultaneously influenced the entire spectrum of liquid
commodities, eventually spurring the transformation of commodity markets
from a collection of idiosyncratic markets into a more cohesive
collection akin to an asset class. In this context, interconnectedness
emerges as a crucial aspect that warrants thorough examination.\\
We strive to address this issue by first examining the average returns
pairwise correlations between commodities across the entire
cross-section of the commodity assets considered in the study as well as
by country, sector and sub-sector.
Table~\ref{tbl-correlations-inner-periods} shows the results for the
four periods of interest with the results reported for the entire period
as well as for periods of \textbf{backwardation} and \textbf{contango}.
For the entire sample the average pairwise correlation jumps from 0.07
in the first period to 0.16 over financialisation. This finding
indicates higher interdependence within the full cross-section of
commodities starting with the financialisation period, consistent with
the empirical findings in \citet{fryMcKibbin_evolution_2023} and
\citet{mayer_financialization_2017} as well as with the theoretical
model of \citet{basak_model_2016}. The sector analysis sheds light on
the driving force behind these changes. The US metals see the sharpest
increase in pairwise correlations going from 0.22 over the past period
to 0.56 over financialisation while the three other sectors see a much
smaller percentage increase during financialisation relative to the past
period. These results thus suggest that the US metals sector was the
driving force behind increased average pairwise correlations during
financialisation, a new observation to the best of our knowledge. The
pattern of correlations is very similar across phases of
\textbf{backwardation} and \textbf{contango} for all the sectors and
sub-sectors over the past and financialisation periods. In this context,
the ontological consequences of financialisation are better observed
from an investment perspective as illustrated in
Table~\ref{tbl-correlations-cross-periods} where we observe average
pairwise correlation coefficients between equally weighted portfolios of
US traded commodity assets grouped by sub-sector\footnote{A more
  detailed breakdown of country, sector and subsector average pairwise
  correlations is provided in the attached
  \href{http://bautheac.duckdns.org:3838/bautheac/co-movement/}{Internet
  Appendix}.}. Although modest in magnitude, these exhibit a telling
pattern, mirroring that shown by the individual commodity returns in
Table~\ref{tbl-stats-regime-tests}, with a substantially higher average
coefficient over \textbf{backwardation} phases in the first period (0.12
vs.~0.03) and the pattern reversed over financilisation with an average
coefficient substantially higher in \textbf{contango} (0.12 vs.~0.02).\\
We break down Table~\ref{tbl-correlations-inner-periods}'s average
correlations by year in Table~\ref{tbl-correlations-inner-years}. For
the full sample of assets, the average pairwise correlations remain in
the 0.06--0.08 range from 1998 through 2003, before suddenly increasing
to 0.12 from 2003 to 2004. This suggests 2004 as a distinct break point
for the onset of financialisation, as noted in several studies
\citep{baker_financialization_2021, tang_index_2012}. We further observe
that the US metals appear to have played a major role in this as the
increase seems driven by the sector which average pairwise correlation
more than doubles (0.49 vs 0.22) over the period.

@@

\textbf{backwardation}\footnote{A breakdown of the performance by regime
  is provided in the attached
  \href{http://bautheac.duckdns.org:3838/bautheac/co-movement/}{Internet
  Appendix}.}. The onset of financialisation leads to a substantial
improvement in the performance of the CHP factor with its average
\(R^{2}\), albeit still small at 5.4\%, almost tripling from the past to
the financialisation period in contrast to the other factors that show
weaker dynamics with a slight decrease and a much smaller increase for
the open interest and term structure factors respectively. Here again
the increase in pricing performance for the CHP factor is driven by the
metals sector, particularly the precious sub-sector with the average
\(R^{2}\) for the group jumping to almost 26\% during financialisation
from 1.2\% over the past period. The corresponding figures of 22\% and
1.1\% for the entire US metals sector stand in sharp contrast with the
agricultural sector, which exhibits no notable change, and the energy
sector, where pricing performance drops substantially from 9.2\% down to
1.6\%. The improvement in the CHP factor's pricing performance
remarkably comes predominantly from the long leg (low CHP assets:
Keynesian ``normal backwardation'') of the factor\footnote{A breakdown
  of the performance by factor leg is provided in the attached
  \href{http://bautheac.duckdns.org:3838/bautheac/co-movement/}{Internet
  Appendix}.} with the corresponding mimicking portfolio's performance
itself driven by the metals, the precious sub-sector in particular.
Interestingly, although the term structure factor itself only sees a
very modest increase, its short leg (low roll-yield assets: Working's
backwardation) shows a sizeable rise in pricing performance, yet again
driven the metals sector. The pattern of results for both factors
extends to the GB traded assets and suggest a convergence of Keneysian
and Working's term structure perspective over the financialisation
period.\\

@@

The crisis period (late 2008 to mid-2013) shows a contrasting pattern of
results. Returns fall sharply relative to the financialisation period
(Table~\ref{tbl-stats-regime-tests}) with eighteen US commodities along
with all six GB traded metals showing lower mean returns. Returns for
equally weighted portfolios of both sets of assets fall down to 9.4\%
and 5.3\% from 15.9\% and 21.8\% respectively. Returns are mostly higher
during phases of \textbf{backwardation} relative to \textbf{contango}
with fifteen US assets and five of the GB traded metals showing higher
mean returns in \textbf{backwardation} and the pattern extending to
equally weighted portfolios of the two sets of assets. Volatility is
moderately higher for both sets with fifteen US commodities and three GB
metals showing higher figures and interestingly eminently higher over
phases of \textbf{contango}; out of the US complex only Sugar-\#11
(IFUS) and none of the GB traded metals show higher volatility figures
over \textbf{backwardation} phases. The symmetry between the dynamics of
CHP and \textbf{CHP}'s explanatory power on individual commodity returns
breaks (Table~\ref{tbl-time-reg-US-CHP}) with both now seemingly
converging; the US cross-sectional average \(R^{2}\) at the individual
level bouncing back up (22.5\% vs 20.8\%) while that at the aggregate
level rises further (14.7\% vs 7.1\%).

@@

The returns pull-back continues into the post-crisis period
(Table~\ref{tbl-stats-regime-tests}) with twenty-two US and four GB
traded assets showing lower mean returns relative to the crisis while
the same number of US commodities and all GB metals show lower figures
relative to the financialisation period; mean returns for equally
weighted portfolios of US and GB assets drop down to -0.2\% and 2.7\%
respectively. Interestingly volatility also recedes substantially with
twenty US assets and all GB metals showing lower volatility figures
relative to both the crisis and financialisation periods suggesting that
investment outflows from the commodity futures complex have had a lower
impact on volatility than the corresponding inflows witnessed over the
financialisation period. The figures for the equally weighted portfolios
of US and GB assets drop to 9.9\% and 16.2\% respectively, very close to
pre-financialisation levels.\\
The pattern of reversion toward the pre-financialisation Keynesian
paradigm initiated during the crisis continues stronger into the
post-crisis period with mean returns higher during phases of
\textbf{backwardation} relative to \textbf{contango}, more so than over
the crisis period with seventeen US commodities showing higher mean
returns over \textbf{backwardation} phases. The difference
(Table~\ref{tbl-stats-regime-tests}) is significant at the 5\% level for
Lumber (XCME) and at the 10\% level for Palladium (XNYM) while no
difference is significant for the assets showing higher figures over
phases of \textbf{contango}. The pattern is more pronounced for the GB
traded subset with all six assets showing higher mean returns in
\textbf{backwardation} with the difference significant at the 1\% level
for Nickel-primary (XLME), 5\% level for Aluminium (XLME) and 10\% level
for Zinc (XLME). It also extends to equally weighted portfolios of US
and GB traded assets that both show higher mean returns during
\textbf{backwardation} with the difference significant at the 5\% level
for the GB portfolio and the figure close to the 10\% level significance
threshold for the portfolio of US assets. Volatilities are overall
higher in \textbf{contango} although less so than over the previous
period with seventeen US commodities showing higher figures over
\textbf{contango} against twenty-three in the crisis period. The pattern
extends to three of the six GB traded metals with the difference
significant at the 1\% level for Copper (XLME) and Tin (XLME) and at the
5\% level for Zinc (XLME) while differences are not significant for
Aluminium (XLME), Lead (XLME) and Nickel-primary (XLME) that exhibit
higher volatility over \textbf{backwardation} phases. Equally weighted
portfolios of US and GB traded assets on the other hand both show higher
volatility in \textbf{contango} with the difference significant at the
1\% level for both. Over the financialisation period in contrast
volatilities were predominantly higher in \textbf{backwardation} for US
traded assets with fourteen commodities showing higher volatility
figures in these phases; the post-crisis pattern of results for
volatilities thus likewise shows signs of reversion to
pre-financialisation standards. After showing signs of convergence in
the crisis period the dynamics of the explanatory power of CHP and
\textbf{CHP} part ways (Table~\ref{tbl-time-reg-US-CHP}), seemingly
reverting to the Keynesian paradigm observed pre-financialisation,
further suggesting a return to legacy commodity specific dynamics. The
recovery initiated in the crisis period for that of CHP is confirmed
with a stronger increase, 27.2\% vs.~22.5\%, strongest for the metals
sector, 37.3\% vs.~23.7\%, while the energy sector shows the opposite
pattern: 11\% vs.~16\%. In contrast, after increasing stronger in the
crisis, that for \textbf{CHP} falls overall post-crisis, 7.5\% vs
14.7\%, with the decrease strongest for the energy sector in a relative
sense: 2.5\% vs.~14.8\%.

@@

\begin{longtable}[t]{>{}l>{}r>{}r>{}r>{}r>{}r}

\caption{\label{tbl-stats-regime-tests}This table shows mean returns and
volatilities for the individual commodities as well as for two equally
weighted portfolios of US and GB traded assets across the four periods.
For each period, estimates are computed independently over the whole
period (figure shown) as well as over phases of \textbf{backwardation}
(b) and \textbf{contango} (c) with corresponding comparison results
reported in brackets. The table further shows the results of t-tests of
significance for the mean returns computed over the whole periods as
well as t-tests of difference between these when computed over the two
phases independently along with F-tests of difference for the
corresponding variances. Values significant at the 1\%, 5\% and 10\%
level are marked with ***, ** and * respectively. The estimate values
per regime as well as the p-values for the corresponding difference
tests are available in the attached
\href{http://bautheac.duckdns.org:3838/bautheac/co-movement/}{Internet
Appendix}. \textbf{Backwardation} and \textbf{contango} regimes as well
as period definitions are discussed in Section~\ref{sec-methods} while
the results are discussed in Section~\ref{sec-results}.}

@@

\textcolor{black}{\textbf{asset}} & \textcolor{black}{\textbf{estimate}} & \textcolor{black}{\textbf{past}} & \textcolor{black}{\textbf{financialisation}} & \textcolor{black}{\textbf{crisis}} & \textcolor{black}{\textbf{post-crisis}}\\

@@

\textbf{Corn-\#2 yellow (XCBT)} & \textbf{mean} & \textcolor[HTML]{4285f4}{\textbf{3.48\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{21.37\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{6.89\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{-4.51\% (b<c)}}\\
\textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{\textbf{23\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{29.54\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{36.15\% (b<c***)}} & \textcolor[HTML]{4285f4}{\textbf{21.8\% (b<c*)}}\\
\textbf{Oats (XCBT)} & \textbf{mean} & \textcolor[HTML]{4285f4}{\textbf{5.49\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{21.62\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{8.89\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{-0.87\% (b>c)}}\\
\textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{\textbf{32.22\% (b<c***)}} & \textcolor[HTML]{4285f4}{\textbf{31.04\% (b>c***)}} & \textcolor[HTML]{4285f4}{\textbf{34.67\% (b<c***)}} & \textcolor[HTML]{4285f4}{\textbf{29.5\% (b>c***)}}\\
\textbf{Soybean meal (XCBT)} & \textbf{mean} & \textcolor[HTML]{4285f4}{\textbf{7.29\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{12.06\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{8.12\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{-0.33\% (b<c)}}\\

@@

\textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{\textbf{26.34\% (b>c***)}} & \textcolor[HTML]{4285f4}{\textbf{32.55\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{31.78\% (b<c***)}} & \textcolor[HTML]{4285f4}{\textbf{25.11\% (b<c)}}\\
\textbf{Soybean oil (XCBT)} & \textbf{mean} & \textcolor[HTML]{4285f4}{\textbf{6.21\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{15.57\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{4.04\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{-7.39\% (b<c)}}\\
\textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{\textbf{21.43\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{28.35\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{25.75\% (b<c***)}} & \textcolor[HTML]{4285f4}{\textbf{18.95\% (b<c**)}}\\
\textbf{Soybeans (XCBT)} & \textbf{mean} & \textcolor[HTML]{4285f4}{\textbf{6.89\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{13.36\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{6.19\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{-3.82\% (b<c)}}\\
\textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{\textbf{21.47\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{30.04\% (b>c***)}} & \textcolor[HTML]{4285f4}{\textbf{28\% (b<c***)}} & \textcolor[HTML]{4285f4}{\textbf{21.22\% (b>c*)}}\\

@@

\textbf{Wheat-SRW (XCBT)} & \textbf{mean} & \textcolor[HTML]{4285f4}{\textbf{5.07\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{17.63\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{6.68\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{-1.58\% (b>c)}}\\
\textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{\textbf{24.93\% (b>c**)}} & \textcolor[HTML]{4285f4}{\textbf{33.03\% (b>c***)}} & \textcolor[HTML]{4285f4}{\textbf{37.62\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{26.58\% (b<c)}}\\
\textbf{Cattle-feeder (XCME)} & \textbf{mean} & \textcolor[HTML]{4285f4}{\textbf{0.39\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{7.43\% (b>c*)}} & \textcolor[HTML]{4285f4}{\textbf{7.13\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{1.87\% (b>c)}}\\
\textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{\textbf{13.23\% (b<c*)}} & \textcolor[HTML]{4285f4}{\textbf{15.16\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{15.24\% (b<c***)}} & \textcolor[HTML]{4285f4}{\textbf{18.69\% (b>c***)}}\\
\textbf{Cattle-live (XCME)} & \textbf{mean} & \textcolor[HTML]{4285f4}{\textbf{3.45\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{8.77\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{4.13\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{1.92\% (b<c)}}\\

@@

\textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{\textbf{17.1\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{17.43\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{16.32\% (b<c***)}} & \textcolor[HTML]{4285f4}{\textbf{18.3\% (b>c***)}}\\
\textbf{Lean hogs (XCME)} & \textbf{mean} & \textcolor[HTML]{4285f4}{\textbf{0.73\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{8.4\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{12.69\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{-1.09\% (b>c)}}\\
\textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{\textbf{38.33\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{31.89\% (b>c***)}} & \textcolor[HTML]{4285f4}{\textbf{30.37\% (b<c***)}} & \textcolor[HTML]{4285f4}{\textbf{36.94\% (b<c*)}}\\
\textbf{Cocoa (IFUS)} & \textbf{mean} & \textcolor[HTML]{4285f4}{\textbf{3.66\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{16.35\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{1.39\% (b<c)}} & \textcolor[HTML]{4285f4}{\textbf{3.65\% (b<c)}}\\
\textbf{} & \textbf{volatility} & \textcolor[HTML]{4285f4}{\textbf{32.66\% (b>c)}} & \textcolor[HTML]{4285f4}{\textbf{29.99\% (b>c*)}} & \textcolor[HTML]{4285f4}{\textbf{31.03\% (b<c***)}} & \textcolor[HTML]{4285f4}{\textbf{24.69\% (b>c**)}}\\
